---
title: "chapter2_2"
author: "Nikunj Goel"
date: "9/21/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(astsa)
```

## 2.6

Judging with our eyes, the time series for gtemp appears non-stationary. 
The mean is non-constant and there is clearly an upward trend. 
The variance appears to be pretty consistent however.

We can further check this through the acf2() function. If stationary, the ACF/PACF plots will show a quick drop-off in correlation after a small amount of lag between points.

```{r cars}
acf2(gtemp)
```
The dotted blue line details significance threshold for each lag. Clearly this data is non-stationary as a high number of previous observations are correlated with future values.

## 2.7



## 2.11

### part a)
```{r}
set.seed(123)
t = 1:500
x = rnorm(t)
acf1(x,20)
```

### part b)

```{r}
set.seed(123)
t=1:50
x = rnorm(t)
acf1(x,20)
```

## 2.12

### part a)

```{r}
w = rnorm(500)
x = filter(w,sides=2,filter=rep(1/3,3))
acf1(x,20)
```

### part b)

```{r}
w = rnorm(50)
x = filter(w,sides=2,filter=rep(1/3,3))
acf1(x,20)
```


## 2.13

```{r}
set.seed(123)
w = rnorm(250+50)
x = filter(w,filter=c(1.5,-0.75),method="recursive")[-(1:50)]
acf1(x,50)
```

## 2.15

```{r}
set.seed(123)
x = sample(c(-2,2), 11,replace=TRUE)
y = 5 + filter(x,sides = 1,filter = c(1,-0.5))[-1]
acf(y, lag.max=4)
```