---
title: "chapter5"
author: "Nikunj Goel"
date: "10/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(astsa)
library(ggplot2)
```

## Problem 5.2

Plotting the *GDP* data.

```{r}
tsplot(gdp)
```

Plotting the *acf* of the data.

```{r}
acf1(gdp)
```

Applying transformation over data

```{r}
tsplot(diff(log(gdp)), ylab="GDP tranformed data", col=4)
abline(h=mean(diff(log(gdp))), col=6)
```
Considering the difference on the logged data for now.

plotting the **acf** and **pacf** plot to determine the **p** and **q** values for the ARIMA model.

```{r}
acf2(diff(log(gdp)))
```
Inspecting the sample **ACF** and **PACF** it feels like there are 3 possibilities. \
1) AR(1) model as PACF looks like cutting off after lag 1 and ACF tails off \
2) MA(2) model as ACF looks like cutting off after lag 2 and PACF tails off \
3) ARMA(1,0,2) on diff(log(gdp)) or ARIMA(1,1,2) on log(gdp).
\
Trying to fit the model over the data

**AR(1)**
```{r}
sarima(log(gdp),p=1,d=1,q=0)
```
**MA(2)**
```{r}
sarima(log(gdp),p=0,d=1,q=2)
```
**ARIMA(1,1,2)**
```{r}
sarima(log(gdp),p=1,d=1,q=2)
```
In this **ARIMA(1,1,2)** has insignificant values as shown by p.value in model fit.\
Standardized residuals in all the models shows no pattern. \
ACF of the residuals shows no apparent departure from the model assumption apart from AR(1). \
P-Value plot shows values exceeding 0.05, we can feel comfortable not rejecting the null hypothesis that the residuals are white apart from AR(1). \

Let's try higher dimension model to see if the value changes

```{r}
sarima(log(gdp),p=2,d=1,q=2)
sarima(log(gdp),p=1,d=1,q=3)
```
Values doesn't change much for higher models. \

Looking at the **AIC**, **AICc** and **BIC** MA(2) is the better choice model of the 3.

## Problem 5.4

First lets see the time series

```{r}
tsplot(gtemp_land)
```
We can see a clear trend, let’s look at the ACF to confirm this,

```{r}
acf(gtemp_land)
```
The slow decay of the ACF suggest us to take a difference, we do that now.

```{r}
tsplot(diff(gtemp_land))
abline(h=mean(diff(gtemp_land)))
```
Looks much better. \

The next step is to plot the ACF and PACF of this differenced gtemp, and establish orders for the ARIMA model.

```{r}
par(mfrow = c(2,1))
acf(diff(gtemp_land))
pacf(diff(gtemp_land))
```
From the above we can see that the ACF appears to cut off at lag 1 and the PACF tails off. This is suggesting the model MA(1) for the difference temp, or ARIMA(0, 1,1) for the gtemp_land. Let use this and fit the model.

```{r}
par(mfrow = c(1,1))
sarima(gtemp_land, 0, 1, 1)
```
Diagnoses \

1) Standardized Residuals: They have to apparent pattern, and so we can conclude tha they resemble white noise. \

2) ACF of Residuals: Much like the Residuals Vs Time plot the ACF of these appears to resemble the acf of white noise and so we are happy. \

3) QQ Plot: The qqplot is a great fit, even towards the tails, the divergence from the theoretical quantiles is minimal. \

Trying more complex models

```{r}
sarima(gtemp_land, 0, 1, 2)
sarima(gtemp_land, 1, 1, 1)
```
Seeing the **AIC** and **AICc**, the model MA(1) or ARIMA(0,1,1) is best.

Forecast

```{r}
sarima.for(gtemp, 10, 0, 1, 1)
```

## Problem 5.11

Plotting the data

```{r}
tsplot(birth)
```

Trying transformations over the data

```{r}
tsplot(diff(diff(birth,12)))
abline(h = mean(diff(diff(birth,12))))
```
Plotting the ACF and PACF of the data

```{r}
acf2(diff(diff(birth,12)))
```
Seasons - It appears that at the seasons, the ACF is cutting off at lag 1s (s=12) whereas the PACF is tailing off at lags 1s,2s,3s,4s. These results simply imply an SMA(1)

Non Seasonal - Inpecting the sample ACF and PACF at the first few lags, it appears as though the ACF cuts off at  lag = 1, whereas the PACF is tailing off. This suggests an MA(1) within the seasons, p=0 and q=1.

```{r}
sarima(birth,p=0,d=1,q=1,P=0,D=1,Q=1,S=12,no.constant = TRUE)
```
p-values of the residuals are all below 0.05 and thus we cannot say it to be white.\
ARIMA(0,1,1) x (0,1,1)_12 does'nt give good results.

let's try for \
ARIMA(0,1,2) x (0,1,1)_12

```{r}
sarima(birth,p=0,d=1,q=2,P=0,D=1,Q=1,S=12,no.constant = TRUE)
sarima(birth,p=1,d=1,q=1,P=0,D=1,Q=1,S=12,no.constant = TRUE)
```

P-values of both the models are decent enough to consider it white but the **AIC** and **AICc** values indicate \
ARIMA(1,1,1) x (0,1,1)_12 as better model.

```{r}
sarima.for(birth,12,1,1,1,0,1,1,12,no.constant = TRUE)
```

## Problem 5.14

Trying to plot the **CPG**

### Part a)

```{r}
tsplot(cpg)
```
The trend of the values are decreasing continously and almost tending to zero when seen for such a long duration.

### Part b)

```{r}
logcpg = log(cpg)
modlm = lm(logcpg ~ time(cpg),na.action = NULL)
summary(modlm)
tsplot(modlm$fitted.values) 
lines(logcpg)
```

The fitted line is nearly perfect to the logged data.
Let $log(c_t) = \beta t$ then \
$c_t = e^{\beta t}$

### Part c)


```{r}
plot(resid(modlm))
acf2(resid(modlm))
```

By inspecting the acf plot of residuals, I can say that the residuals are not white.


### Part d)

PACF lags at lag=1 and ACF tails off from **Part c)**. \
Let's try AR(1) model.

```{r}
reg2 = sarima(logcpg,1,0,0,xreg= time(cpg))
```

After inspecting the residuals and ACF it looks fairly white.
